\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-05-06 forwards-and-backwards.tex
Update on Overleaf.

\comment{TODO: talk about $N \to \infty$ rather than $\theta \to \infty$.}

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    Recall that we have defined the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Normal
    with mean $\meanq/\theta$ and covariance matrix $\covq/\theta$.
    This implies that if we define, for $f : \IR^d \to \IR$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } \theta \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    $$
    \DG^* f(x)
        =
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq_i) .
    $$
\end{definition}

\paragraph{Remark:}
An equivalent way to describe the model
would be to say that an individual at $x$ gives birth at rate
$$
    \gamma(x, \smooth{\gamma} \eta(x))
    \int r(x, \smooth{r} \eta(y)) q(x, dy) ,
$$
and that offspring disperse according to the kernel
$$
    q_\theta^\mathfrak{m}(x, dy)
    :=
    \frac{
        r(y, \smooth{r} \eta(y)) q_\theta(x, dy)
    }{
        \int r(x, \smooth{r} \eta(z)) q_\theta(x, dz)
    } .
$$
Clearly, the random walk driven by this dispersal kernel
should be biased towards regions of higher establishment probability.
For comparison to future results,
it is interesting to write down the limiting generator:
\begin{align*}
    \lim_{\theta \to \infty}
    \theta \int (f(y) - f(x)) q_\theta^\mathfrak{m}(x, dy)
    &=
    \frac{
        \DG\left[ f(\cdot) r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
        - 
        f(\cdot) \DG\left[ r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
    }{
        r(x, \smooth{r} \eta^{\theta}(x))
    } .
\end{align*}
In the simplest case of unbiased, isotropic dispersal,
$\meanq = 0$ and $\covq = I$, so $\DG = \Delta$,
and this is equal to
\begin{align*}
    \Delta f(x) + 2 \grad f(x) \cdot \grad \log r(\cdot, \smooth{r} \eta(\cdot))(x) .
\end{align*}
However, we will see that following a lineage back through the population
has a different behavior yet again.

@@

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^\theta_t)_{t \geq 0}$
    be as defined in Definition \ref{def:model_setup}
    and assume that as $\theta \to \infty$, $N \to \infty$
    in such a way that $\theta/N \to \alpha$.
    (However, interaction distances $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$
    remain constant.)

    The sequence of stochastic processes
    $(\eta^\theta_t)_{t \geq 0}$
    converges along subsequences in distribution as $\theta \to \infty$
    to a measure-valued process $(\eta_t)_{t \geq 0}$
    which satisfies that, for every $f:\mathbb{R}^d \rightarrow \mathbb{R}$,
    \begin{align} \label{eqn:mgale_problem}
        M_t
        &:=
        \langle f(x), \eta_t(dx) \rangle
        -
        \langle f(x), \eta_0(dx) \rangle
        \\ & \qquad
        -
        \int_0^t \big\langle
            \gamma(x, \smooth{\gamma} \eta_{s}(x))
            \mathcal{B}\left(
                f(\cdot) r(\cdot, \smooth{r} \eta_{s}(\cdot))
            \right)(x)
            +
            f(x)
            F(x, \smooth{F} \eta_{s}(x)),
            \eta_{s}(dx)
        \big\rangle ds,
    \end{align}
    is a martingale.
    The quadratic variation of $M_t$, denoted $\langle M \rangle_t$, is given by 
    \begin{align} \label{eqn:mgale_variation}
        \langle M \rangle_t
        =
        \alpha
        \int_0^t
        \big\langle
            \gamma\left( x, \smooth{\gamma} \eta_{s}(x) \right)
            r\left(x, \smooth{r} \eta_{s}(x) \right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    In other words, the limit is a measure-valued process with non-local interactions.
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

\begin{corollary}[Superprocess limits] \label{cor:superprocess_uniqueness}
    In the context of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations \eqref{eqn:mgale_problem} and \eqref{eqn:mgale_variation}
    has a unique solution,
    then $(\eta^\theta(t))_{t \ge 0}$ converges to that solution
    as $\theta \to \infty$.
    \comment{
        Is this what we want to say here? Could be said better.
        TODO: see what Dawson \& Li say about uniqueness
        so we have at least some conditions under which this happens.
    }
\end{corollary}

@@

Theorem \ref{thm:nonlocal_convergence} combined with Proposition \ref{prop:nonlocal_to_local}
imply that we can take the $\theta \to \infty$ limit
followed by $\epsilon \to 0$
to obtain solutions to the PDE \eqref{eqn:PDE}.
However, it is of substantial interest to know whether
we can take those two limit simultaneously.
The general case seems difficult,
but we can prove such ``diagonal'' convergence in the following situation.

\begin{theorem}[Convergence to a PDE]
    \label{thm:local_convergence}
    Let $(\eta^\theta_t)_{t \geq 0}$
    be as defined in Definition \ref{def:model_setup}
    and assume that as $\theta \to \infty$, $N \to \infty$
    and $\epsilon_r < \epsilon_\gamma = \epsilon_F = \epsilon \to 0$
    in such a way that $\theta/N \to 0$
    and
    \comment{TODO: some conditions on $\epsilon$ and $r$ and $\gamma$ and $F$}.
    Then the sequence of stochastic processes $(\eta^\theta_t)_{t \ge 0}$
    converges \comment{in some sense}
    to a measure-valued process with a density $\varphi(t, x)$
    that solves
    \begin{align}
        \partial_t \varphi(t, x)
        &=
        \text{whatever the final form is here} .
    \end{align}
\end{theorem}

@@

Equation~\eqref{eqn:rewritten mean measure}) then converges to
\begin{equation} \label{limit of mean measure equation}
\int \gamma(x,\eta)
\DG \big(f(\cdot)r(\cdot,\eta)\big)(x)
\eta(dx)
+
\int f(x)
F(x,\eta)
\eta(dx) .
\end{equation}
More generally, we expect everything to work
if the remaining parameters ($r$, $\gamma$, and $F$)
depended on $\theta$, but converged as $\theta \to \infty$.

@@

We now look at the second order term.
An individual at location $x$ gives birth 
to a surviving offspring at $y$ at rate
$$
\gamma(x,\eta) r(y,\eta) q_{\theta}(x, dy) ,
$$
and since this increments $\langle f, \eta \rangle$ by $f(y) / N$,
the contribution to the quadratic variation from birth events,
which occur at rate $\theta$ per individual 
(so, rate $N\theta |\eta|$ overall), is
$$
\langle
    N \theta \gamma(x,\eta)
    \int \frac{1}{N^2} f^2(y) r(y,\eta)
    q_\theta(x,dy) 
    , \eta(dx)
\rangle .
$$
Similarly, the increment in $\langle f, \eta\rangle$ resulting from 
the death of an individual at $x$ is $f(x)/N$, and so combining with the 
above, the second order term in the generator takes the form
$$
F''(\langle f,\eta\rangle)
\frac{1}{2} N \theta
\left\{
    \langle
        \gamma(x,\eta)
        \int \frac{1}{N^2}f^2(y)r(y,\eta)q_\theta(x,dy) 
    , \eta(dx)\rangle
    +
    \langle
        \mu_\theta(x,\eta)\frac{1}{N^2}f^2(x) 
    ,\eta(dx)\rangle
\right\} .
$$
The term $\epsilon_{\theta,N}(f, F, \eta)$ will be 
$\bigO(\theta/N^2)$.

@@

If $N=\alpha\theta$, for some positive constant $\alpha$,
the second order term in the generator remains order
one as $\theta\to\infty$, and  
we expect a `generalised superprocess' limit.
Assuming some regularity of the coefficients (in particular of $r$), 
the second order term in the generator will take the form
$$
\frac{1}{\alpha}F''(\langle f,\Xi\rangle)
\langle f^2(x), \gamma(x,\Xi)
r(x,\Xi)\Xi(dx)\rangle.
$$
That is
$$\langle f(x), \Xi_t(dx)\rangle-\langle f(x), \Xi_0(dx)\rangle
-\int_0^t\langle\gamma(x,\Xi_s)\DG\big(r(\cdot,\Xi_s)f(\cdot)\big)(x),
\Xi_s(dx)\rangle ds$$
is a martingale with quadratic variation
$$\frac{2}{\alpha}\int_0^t\langle f^2(x) 
\gamma(x,\Xi_s)r(x,\Xi_s),\Xi_s(dx)\rangle ds.$$

@@

Although our proof of Theorem~\ref{thm:lineages}
uses the explicit representation in terms of the lookdown process,
informal calculations agree with the main result.
Suppose that we are tracking a lineage that is at location $y$ at time $t$.
Looking back through time, the lineage will move at the time that individual was born,
and will jump to the location of the parent of that individual.
Now, the rate at which new individuals are born to parents at $x$ and establish at $y$
is
$$
    \gamma(x, \eta_t) q(x, dy) r(y, \eta_t) .
$$
\comment{TODO: explain this more.}
Happily, we are interested in the case when $\eta$ has a density,
say $\eta_t(dx) = \varphi_t(x) dx$,
so interpreting these rates is not too difficult.
This implies that the probability a randomly chosen individual near $y$
is a new offspring from a parent at $x$ in $[t, t+dt)$ is, informally,
\begin{equation} \label{eqn:informal_rates}
\frac{
    \gamma(x, \eta_t) r(y, \eta_t) \varphi_t(x)
}{
    \varphi_t(y)
} \frac{ q_\theta(x, dy) }{ dy } dx dt .
\end{equation}
Leaving aside questions of whether a lineage can be treated as a randomly chosen individual,
we might then proceed by defining a continuous-time jump process
whose transition rates are given by \eqref{eqn:informal_rates} given $(\varphi_t)_{t=0}^T$.
Let $(L^\theta_s)_{s=0}^T$ be the location of a lineage that moves according to
jump rates~\eqref{eqn:informal_rates},
back through time (i.e., with the substitution that $s = T-t$).
Then, writing $q_\theta(x, y)$ for the density of $q_\theta$,
\begin{align} \label{eqn:lineage_generator_Edt}
    \begin{split}
    &\IE[f(L^\theta_{s+ds}) - f(y) \;|\; L_s = y]
    \\&\qquad 
    =
    ds \int \left(f(x) - f(y)\right)
    \frac{
        \varphi_{T-s}(x) \gamma(x, \eta_{T-s}) r(y, \eta_{T-s})
    }{
        \varphi_{T-s}(y)
    }
    q_\theta(x, y) dx .
    \end{split}
\end{align}
\comment{I'm not sure if at this point in the paper we want to either
(a) write out the next statement; (b) restrict to the $\DG = \Delta$ case,
or (c) leave it as is (as an exercise for the reader, I guess?}
Referring back to Definition~\ref{def:dispersal_generator},
it is not too hard to show that as $\theta \to \infty$,
\begin{align*}
    \theta \int (f(x) - f(y)) q_\theta(x, y) dx
    &\to
    \DG^* f(x)  - f(x) \DG^* 1 .
\end{align*}
Consequently,
\begin{align*}
    &
    \theta \int (f(x) - f(y)) g(x) q_\theta(x, y) dx \\
    &\qquad =
    \theta \int \left\{
        (f(x) g(x) - f(y) g(y)) - f(y) (g(x) - g(y))
    \right\} q_\theta(x, y) dx \\
    &\qquad \to
        \DG^*(fg)(y) - f(y) \DG^* g(y) . 
\end{align*}
Applying this to~\eqref{eqn:lineage_generator_Edt} with $g = \varphi_{T-s} \gamma$,
this suggests that the generator of the limiting process is
\begin{align} \label{eqn:heuristic_lineage_generator}
    \Lgen_s f
    &=
    \frac{r}{\varphi_{T-s}}
    \left\{
        \DG^*(\gamma \varphi_{T-s} f) - f \DG^*(\gamma \varphi_{T-s})
    \right\} .
\end{align}
This agrees with Theorem~\ref{thm:lineages}.

@@

Putting these together,
the level $u$ of an individual at $x$ in $\lp$ evolves according to:
\begin{equation} \label{eqn:dot_u}
    \dot u
    =
    - \frac{\theta}{\lambda} u \left(\lambda - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) 
    +
    \theta \mu(x,\eta) u .
\end{equation}
The factor of $\theta$ appears here because we have rescaled time.
We shall write 
$$
    b(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    -
    \mu(x,\eta)
    \right) ,
$$
which captures the local net difference between reproduction and death,
i.e., the deviation from criticality of the branching mechanism.
Recall from equation \eqref{eqn:mu_defn} that
$F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu(x,\eta))$,
and so we will have that
\begin{align} \label{eqn:b_limit}
\begin{split}
b(x, \eta)
&=
    \theta \gamma(x, \eta) \int_{\IR^d} \left( r(y, \eta) - r(x, \eta) \right) q(x, dy)
    + F(x, \eta) \\
&\to
    \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta) \qquad \text{as } \theta \to \infty .
\end{split}
\end{align}

@@

Theorem A.10 in \cite{kurtz/rodrigues:2011}
states that convergence of $(\eta^{\theta}_t)_{t \geq 0}$
to a conditionally Poisson system $(\xi_t)_{t \geq 0}$
with Cox measure $\Xi_t \times \Lambda$
is equivalent to convergence of
$(\eta^{\theta}_t)_{t \geq 0}$ to $(\Xi_t)_{t \geq 0}$.
As a result, it suffices for us to focus on
tightness of $(\eta^{\theta}_t)_{t \geq 0}$
and spatial convergence of $(\eta^{\theta}_t)_{t \geq 0}$
as $\theta \to \infty$.

@@

\begin{align}
F'(s)=&  -\frac{\epsilon^d_r |\covq|}
        {2\left|\covq s +  \epsilon^2_r I \right|^{3/2}}
        \exp \left(
                    \frac{(x-w-\meanq s)^{T}
                        (\covq s +  \epsilon^2_r I)^{-1}
                        (x-w-\meanq s)}{2}
                        -\frac{|x-w|^2}{2\epsilon^2}
            \right)\\
    &+\frac{\epsilon^d_r }
        {\left|\covq s +  \epsilon^2_r I \right|^{1/2}}
        \exp \left(
                    \frac{(x-w-\meanq s)^{T}
                        (\covq s +  \epsilon^2_r I)^{-1}
                        (x-w-\meanq s)}{2}
                        -\frac{|x-w|^2}{2\epsilon^2}
            \right) \\
    & \qquad \times \frac{1}{2}
                    \bigg\{
                         -\meanq^{T}(\covq s + \epsilon^2_r I)^{-1}(x-w-\meanq s)\\
    & \qquad \qquad     -(x-w-\meanq s)^{T}
                        (\covq s +  \epsilon^2_r I)^{-1}\meanq\\
    & \qquad \qquad     -(x-w-\meanq s)^{T}
                        \covq(\covq s +  \epsilon^2_r I)^{-2}(x-w-\meanq s)                    
                    \bigg\},
\end{align}
which is continuous and bounded near $s=0$.
As a result,
\begin{align}
\left| \theta \int_{\mathbb{R}^d}
                \left(p_{\epsilon^2_r}(y-w)
                        -p_{\epsilon^2_r}(x-w)
                \right)
                q_\theta(x,dy)
        \right|
        \leq&  \theta \left(F\left(1 / \theta \right)-1\right) 
                \times \epsilon_\gamma^d \epsilon_r^{-d}
                p_{\epsilon^2_\gamma}(x-w)\\
         =&  F'(u) 
                \times \epsilon_\gamma^d \epsilon_r^{-d}
                p_{\epsilon^2_\gamma}(x-w)        
\end{align}
for some $u \in [0, 1/\theta]$ by Intermediate Value Theorem. We have therefore established our bound as $F(s)$ has bounded derivative in a neighbourhood around $0$.

@@

\begin{equation}
\limsup_{\theta \to \infty}\mathbb{P}\left[\sup_{s \in [0,T]}\langle 1 ,\eta^{\theta}_{s}\rangle \geq K\right] \leq \frac{C(T)}{K}    
\end{equation}
for any $K > 0$, giving us compact containment condition.

@@

Before we prove convergence, we define our limit $(\eta^{\infty}(t))_{t>0}$ as a solution to a Martingale Problem with generator $\Pgen$.

>>>>

\comment{TODO: talk about $N \to \infty$ rather than $N \to \infty$.}

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    Recall that we have defined the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Normal
    with mean $\meanq/\theta$ and covariance matrix $\covq/\theta$.
    This implies that if we define, for $f : \IR^d \to \IR$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } N \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    $$
    \DG^* f(x)
        =
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq_i) .
    $$
\end{definition}

\paragraph{Remark:}
An equivalent way to describe the model
would be to say that an individual at $x$ gives birth at rate
$$
    \gamma(x, \smooth{\gamma} \eta(x))
    \int r(x, \smooth{r} \eta(y)) q(x, dy) ,
$$
and that offspring disperse according to the kernel
$$
    q_\theta^\mathfrak{m}(x, dy)
    :=
    \frac{
        r(y, \smooth{r} \eta(y)) q_\theta(x, dy)
    }{
        \int r(x, \smooth{r} \eta(z)) q_\theta(x, dz)
    } .
$$
Clearly, the random walk driven by this dispersal kernel
should be biased towards regions of higher establishment probability.
For comparison to future results,
it is interesting to write down the limiting generator:
\begin{align*}
    \lim_{N \to \infty}
    \theta \int (f(y) - f(x)) q_\theta^\mathfrak{m}(x, dy)
    &=
    \frac{
        \DG\left[ f(\cdot) r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
        - 
        f(\cdot) \DG\left[ r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
    }{
        r(x, \smooth{r} \eta^{\theta}(x))
    } .
\end{align*}
In the simplest case of unbiased, isotropic dispersal,
$\meanq = 0$ and $\covq = I$, so $\DG = \Delta$,
and this is equal to
\begin{align*}
    \Delta f(x) + 2 \grad f(x) \cdot \grad \log r(\cdot, \smooth{r} \eta(\cdot))(x) .
\end{align*}
However, we will see that following a lineage back through the population
has a different behavior yet again.

@@

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^\theta_t)_{t \geq 0}$
    be as defined in Definition \ref{def:model_setup}
    and assume that as $N \to \infty$, $N \to \infty$
    in such a way that $\theta/N \to \alpha$.
    (However, interaction distances $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$
    remain constant.)

    The sequence of stochastic processes
    $(\eta^\theta_t)_{t \geq 0}$
    converges along subsequences in distribution as $N \to \infty$
    to a measure-valued process $(\eta_t)_{t \geq 0}$
    which satisfies that, for every $f:\mathbb{R}^d \rightarrow \mathbb{R}$,
    \begin{align} \label{eqn:mgale_problem}
        M_t
        &:=
        \langle f(x), \eta_t(dx) \rangle
        -
        \langle f(x), \eta_0(dx) \rangle
        \\ & \qquad
        -
        \int_0^t \big\langle
            \gamma(x, \smooth{\gamma} \eta_{s}(x))
            \mathcal{B}\left(
                f(\cdot) r(\cdot, \smooth{r} \eta_{s}(\cdot))
            \right)(x)
            +
            f(x)
            F(x, \smooth{F} \eta_{s}(x)),
            \eta_{s}(dx)
        \big\rangle ds,
    \end{align}
    is a martingale.
    The quadratic variation of $M_t$, denoted $\langle M \rangle_t$, is given by 
    \begin{align} \label{eqn:mgale_variation}
        \langle M \rangle_t
        =
        \alpha
        \int_0^t
        \big\langle
            \left(\gamma\left( x,  \eta_{s} \right)
            r\left(x, \eta_{s} \right)+\mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    In other words, the limit is a measure-valued process with non-local interactions.
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

\begin{corollary}[Superprocess limits] \label{cor:superprocess_uniqueness}
    In the context of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations \eqref{eqn:mgale_problem} and \eqref{eqn:mgale_variation}
    has a unique solution,
    then $(\eta^\theta(t))_{t \ge 0}$ converges to that solution
    as $N \to \infty$.
    \comment{
        Is this what we want to say here? Could be said better.
        TODO: see what Dawson \& Li say about uniqueness
        so we have at least some conditions under which this happens.
    }
\end{corollary}

@@

Theorem \ref{thm:nonlocal_convergence} combined with Proposition \ref{prop:nonlocal_to_local}
imply that we can take the $N \to \infty$ limit
followed by $\epsilon \to 0$
to obtain solutions to the PDE \eqref{eqn:PDE}.
However, it is of substantial interest to know whether
we can take those two limit simultaneously.
The general case seems difficult,
but we can prove such ``diagonal'' convergence in the following situation.

\begin{theorem}[Convergence to a PDE]
    \label{thm:local_convergence}
    Let $(\eta^\theta_t)_{t \geq 0}$
    be as defined in Definition \ref{def:model_setup}
    and assume that as $N \to \infty$, $N \to \infty$
    and $\epsilon_r < \epsilon_\gamma = \epsilon_F = \epsilon \to 0$
    in such a way that $\theta/N \to 0$
    and
    \comment{TODO: some conditions on $\epsilon$ and $r$ and $\gamma$ and $F$}.
    Then the sequence of stochastic processes $(\eta^\theta_t)_{t \ge 0}$
    converges \comment{in some sense}
    to a measure-valued process with a density $\varphi(t, x)$
    that solves
    \begin{align}
        \partial_t \varphi(t, x)
        &=
        \text{whatever the final form is here} .
    \end{align}
\end{theorem}

@@

Equation~\eqref{eqn:rewritten mean measure}) then converges to
\begin{equation} \label{limit of mean measure equation}
\int \gamma(x,\eta)
\DG \big(f(\cdot)r(\cdot,\eta)\big)(x)
\eta(dx)
+
\int f(x)
F(x,\eta)
\eta(dx) .
\end{equation}
More generally, we expect everything to work
if the remaining parameters ($r$, $\gamma$, and $F$)
depended on $\theta$, but converged as $N \to \infty$.

@@

We now look at the second order term.
An individual at location $x$ gives birth 
to a surviving offspring at $y$ at rate
$$
\gamma(x,\eta) r(y,\eta) q_{\theta}(x, dy) ,
$$
and since this increments $\langle f, \eta \rangle$ by $f(y) / N$,
the contribution to the quadratic variation from birth events,
which occur at rate $\theta$ per individual 
(so, rate $N\theta |\eta|$ overall), is
$$
\langle
    N \theta \gamma(x,\eta)
    \int \frac{1}{N^2} f^2(y) r(y,\eta)
    q_\theta(x,dy) 
    , \eta(dx)
\rangle .
$$
Similarly, the increment in $\langle f, \eta\rangle$ resulting from 
the death of an individual at $x$ is $f(x)/N$, and so combining with the 
above, the second order term in the generator takes the form
$$
F''(\langle f,\eta\rangle)
\frac{1}{2} N \theta
\left\{
    \langle
        \gamma(x,\eta)
        \int \frac{1}{N^2}f^2(y)r(y,\eta)q_\theta(x,dy) 
    , \eta(dx)\rangle
    +
    \langle
        \mu(x,\eta)\frac{1}{N^2}f^2(x) 
    ,\eta(dx)\rangle
\right\} .
$$
The term $\epsilon_{\theta,N}(f, F, \eta)$ will be 
$\bigO(\theta/N^2)$.

@@

If $N=\alpha\theta$, for some positive constant $\alpha$,
the second order term in the generator remains order
one as $\theta\to\infty$, and  
we expect a `generalised superprocess' limit.
Assuming some regularity of the coefficients (in particular of $r$), 
the second order term in the generator will take the form
$$
\frac{1}{\alpha}F''(\langle f,\Xi\rangle)
\langle f^2(x), \gamma(x,\Xi)
r(x,\Xi)\Xi(dx)\rangle.
$$
That is
$$\langle f(x), \Xi_t(dx)\rangle-\langle f(x), \Xi_0(dx)\rangle
-\int_0^t\langle\gamma(x,\Xi_s)\DG\big(r(\cdot,\Xi_s)f(\cdot)\big)(x),
\Xi_s(dx)\rangle ds$$
is a martingale with quadratic variation
$$\frac{2}{\alpha}\int_0^t\langle f^2(x) 
(\gamma(x,\Xi_s)r(x,\Xi_s)+\mu(x,\Xi_s)),\Xi_s(dx)\rangle ds.$$

@@

Although our proof of Theorem~\ref{thm:lineages}
uses the explicit representation in terms of the lookdown process,
informal calculations agree with the main result.
Suppose that we are tracking a lineage that is at location $y$ at time $t$.
Looking back through time, the lineage will move at the time that individual was born,
and will jump to the location of the parent of that individual.
Now, the rate at which new individuals are born to parents at $x$ and establish at $y$
is
$$
    \gamma(x, \eta_t) q(x, dy) r(y, \eta_t) .
$$
\comment{TODO: explain this more.}
Happily, we are interested in the case when $\eta$ has a density,
say $\eta_t(dx) = \varphi_t(x) dx$,
so interpreting these rates is not too difficult.
This implies that the probability a randomly chosen individual near $y$
is a new offspring from a parent at $x$ in $[t, t+dt)$ is, informally,
\begin{equation} \label{eqn:informal_rates}
\frac{
    \gamma(x, \eta_t) r(y, \eta_t) \varphi_t(x)
}{
    \varphi_t(y)
} \frac{ q_\theta(x, dy) }{ dy } dx dt .
\end{equation}
Leaving aside questions of whether a lineage can be treated as a randomly chosen individual,
we might then proceed by defining a continuous-time jump process
whose transition rates are given by \eqref{eqn:informal_rates} given $(\varphi_t)_{t=0}^T$.
Let $(L^\theta_s)_{s=0}^T$ be the location of a lineage that moves according to
jump rates~\eqref{eqn:informal_rates},
back through time (i.e., with the substitution that $s = T-t$).
Then, writing $q_\theta(x, y)$ for the density of $q_\theta$,
\begin{align} \label{eqn:lineage_generator_Edt}
    \begin{split}
    &\IE[f(L^\theta_{s+ds}) - f(y) \;|\; L_s = y]
    \\&\qquad 
    =
    ds \int \left(f(x) - f(y)\right)
    \frac{
        \varphi_{T-s}(x) \gamma(x, \eta_{T-s}) r(y, \eta_{T-s})
    }{
        \varphi_{T-s}(y)
    }
    q_\theta(x, y) dx .
    \end{split}
\end{align}
\comment{I'm not sure if at this point in the paper we want to either
(a) write out the next statement; (b) restrict to the $\DG = \Delta$ case,
or (c) leave it as is (as an exercise for the reader, I guess?}
Referring back to Definition~\ref{def:dispersal_generator},
it is not too hard to show that as $N \to \infty$,
\begin{align*}
    \theta \int (f(x) - f(y)) q_\theta(x, y) dx
    &\to
    \DG^* f(x)  - f(x) \DG^* 1 .
\end{align*}
Consequently,
\begin{align*}
    &
    \theta \int (f(x) - f(y)) g(x) q_\theta(x, y) dx \\
    &\qquad =
    \theta \int \left\{
        (f(x) g(x) - f(y) g(y)) - f(y) (g(x) - g(y))
    \right\} q_\theta(x, y) dx \\
    &\qquad \to
        \DG^*(fg)(y) - f(y) \DG^* g(y) . 
\end{align*}
Applying this to~\eqref{eqn:lineage_generator_Edt} with $g = \varphi_{T-s} \gamma$,
this suggests that the generator of the limiting process is
\begin{align} \label{eqn:heuristic_lineage_generator}
    \Lgen_s f
    &=
    \frac{r}{\varphi_{T-s}}
    \left\{
        \DG^*(\gamma \varphi_{T-s} f) - f \DG^*(\gamma \varphi_{T-s})
    \right\} .
\end{align}
This agrees with Theorem~\ref{thm:lineages}.

@@

Putting these together,
the level $u$ of an individual at $x$ in $\lp$ evolves according to:
\begin{equation} \label{eqn:dot_u}
    \dot u
    =
    - \frac{\theta}{\lambda} u \left(\lambda - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) 
    +
    \theta \mu(x,\eta) u .
\end{equation}
The factor of $\theta$ appears here because we have rescaled time.
We shall write 
$$
    b(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    -
    \mu(x,\eta)
    \right) ,
$$
which captures the local net difference between reproduction and death,
i.e., the deviation from criticality of the branching mechanism.
Recall from equation \eqref{eqn:mu_defn} that
$F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu(x,\eta))$,
and so we will have that
\begin{align} \label{eqn:b_limit}
\begin{split}
b(x, \eta)
&=
    \theta \gamma(x, \eta) \int_{\IR^d} \left( r(y, \eta) - r(x, \eta) \right) q(x, dy)
    + F(x, \eta) \\
&\to
    \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta) \qquad \text{as } N \to \infty .
\end{split}
\end{align}

@@

Theorem A.10 in \cite{kurtz/rodrigues:2011}
states that convergence of $(\eta^{\theta}_t)_{t \geq 0}$
to a conditionally Poisson system $(\xi_t)_{t \geq 0}$
with Cox measure $\Xi_t \times \Lambda$
is equivalent to convergence of
$(\eta^{\theta}_t)_{t \geq 0}$ to $(\Xi_t)_{t \geq 0}$.
As a result, it suffices for us to focus on
tightness of $(\eta^{\theta}_t)_{t \geq 0}$
and spatial convergence of $(\eta^{\theta}_t)_{t \geq 0}$
as $N \to \infty$.

@@

\begin{align}
F'(s)=&  -\frac{\epsilon^d_r |\covq|}
        {2\left|\covq s +  \epsilon^2_r I \right|^{3/2}}
        \exp \left(
                    \frac{(x-w-\meanq s)^{T}
                        (\covq s +  \epsilon^2_r I)^{-1}
                        (x-w-\meanq s)}{2}
                        -\frac{|x-w|^2}{2\epsilon^2}
            \right)\\
    &+\frac{\epsilon^d_r }
        {\left|\covq s +  \epsilon^2_r I \right|^{1/2}}
        \exp \left(
                    \frac{(x-w-\meanq s)^{T}
                        (\covq s +  \epsilon^2_r I)^{-1}
                        (x-w-\meanq s)}{2}
                        -\frac{|x-w|^2}{2\epsilon^2}
            \right) \\
    & \qquad \times \frac{1}{2}
                    \bigg\{
                         -\meanq^{T}(\covq s + \epsilon^2_r I)^{-1}(x-w-\meanq s)\\
    & \qquad \qquad     -(x-w-\meanq s)^{T}
                        (\covq s +  \epsilon^2_r I)^{-1}\meanq\\
    & \qquad \qquad     -(x-w-\meanq s)^{T}
                        \covq(\covq s +  \epsilon^2_r I)^{-2}(x-w-\meanq s)                    
                    \bigg\},
\end{align}
which is continuous and bounded near $s=0$.
As a result,
\begin{align}
\left| \theta \int_{\mathbb{R}^d}
                \left(p_{\epsilon^2_r}(y-w)
                        -p_{\epsilon^2_r}(x-w)
                \right)
                q_\theta(x,dy)
        \right|
        \leq&  \theta \left(F\left(1 / \theta \right)-1\right) 
                \times \epsilon_\gamma^d \epsilon_r^{-d}
                p_{\epsilon^2_\gamma}(x-w)\\
         =&  F'(u) 
                \times \epsilon_\gamma^d \epsilon_r^{-d}
                p_{\epsilon^2_\gamma}(x-w)        
\end{align}
for some $u \in [0, 1/\theta]$ by Mean Value Theorem. We have therefore established our bound as $F(s)$ has bounded derivative in a neighbourhood around $0$.

@@

\begin{equation}
\limsup_{N \to \infty}\mathbb{P}\left[\sup_{s \in [0,T]}\langle 1 ,\eta^{\theta}_{s}\rangle \geq K\right] \leq \frac{C(T)}{K}    
\end{equation}
for any $K > 0$, giving us compact containment condition.

@@

Before we prove convergence, we define our limit $(\eta^{\infty}(t))_{t>0}$ as a solution to a Martingale Problem with generator $\Pgen^{\infty}$.
